\documentclass[edeposit,fullpage]{uiucthesis07} 

\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{subfigure}
\usepackage{tabularx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[toc,page]{appendix}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions} % library to create annotations 
\usetikzlibrary{shapes.geometric}


\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\Hhybtwid}{\tilde{H}_\text{hyb}}
\newcommand{\Tr}{\text{Tr}}
\newcommand{\ket}[1]{\left|#1\right\rangle}
\newcommand{\mr}[1]{\mathrm{#1}}

\def\prb{\ref@jnl{Phys.~Rev.~B}}        % Physical Review B: Solid State

\bibliographystyle{apsrev}

\begin{document}

\pagenumbering{roman} \thispagestyle{empty} \begin{singlespace}

\begin{center}
\begin{Large}
%YOUR TITLE
STUDY OF CONDENSED MATTER SYSTEMS 
WITH MONTE CARLO SIMULATION
ON HETEROGENEOUS COMPUTING SYSTEMS
\end{Large}

\vspace{2.65in}

A Dissertation\\
\vspace{1.5ex}
Submitted to the Graduate Faculty of the\\
Louisiana State University and\\
Agricultural and Mechanical College\\
in partial fulfillment of the\\
requirements for the degree of\\
Doctor of Philosophy\\
\vspace{1ex}
in\\
\vspace{1ex}
The Department of Physics and Astronomy\\
\vspace{2.65in}
by\\
%Your Name\\
Sheng Feng\\
%B.S.,Institusion?, 200?\\
B.Sc., University of Science and Technology of China, 2010\\

\end{center}
\end{singlespace}
\frontmatter


\begin{dedication}
  ...
\end{dedication}

\chapter*{Acknowledgements}
\markboth{Acknowledgments}{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

\tableofcontents 

\chapter*{Abstract}
\markboth{Abstract}{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

We study the Edwards-Anderson model on a simple cubic lattice with a finite 
constant external field. We implement and optimize a code with CUDA 
architecture for efficient Monte Carlo simulation of spin glass systems on 
Graphic Processing Units. 
We study several traditional indicators, including the Binder ratio and the 
correlation length, as well as a new quantity composed of 
a ratio of susceptibilities at finite momenta.% which was recently proposed to 
%avoid the difficulties of a zero momentum quantity, for capturing the spin glass
% phase transition. 
%Unfortunately, this new indicator is fairly noisy, so 
%a large pool of samples at low temperature and small external field are 
%needed to generate results with sufficiently small statistical error for 
%analysis.  
We confirm previous findings that conventional indicators for the 
spin glass transition do not show any indication of a transition for rather 
low temperatures. 
However, the ratio of spin glass susceptibilities do show crossing behavior, 
albeit a systematic analysis is beyond the reach of the present data. 
%This calls for a more thorough study of the three dimensional Edwards-Anderson 
%model in an external field. 


One of the fundamental challenges of theoretical condensed matter physics is 
the accurate solution of quantum impurity models. By taking expansion in the 
hybridization about an exactly solved local limit, one can formulate an quantum
impurity solver. We implement the hybridization expansion quantum impurity
solver on Intel Xeon Phi accelerators, and aim to apply this approach on the
dynamic Hubbard Models.
%Dynamic Hubbard Models are proposed to reflect the fact that the wavefunction 
%of an electron in an atomic orbital expands when a second electron occupies 
%the orbital. 





\mainmatter
\include{chapters/intro}
\include{chapters/chap1}
\include{chapters/chap2}
\include{chapters/chap3}
\include{chapters/chap4}
%\include{chapters/chap5}
%\include{chapters/cthyb}
\include{chapters/conclusion}
\backmatter
\bibliography{thesis}

\appendix
\chapter*{Appendix: Notes on covariance matrix spectrum}
\include{chapters/appendix_matrix}



\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
